---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval=FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
```

```{r, include=FALSE, eval=TRUE}
# If you actually want to eval all the code here to check that it runs,
# change actually_eval to TRUE and make sure you run this on a Monday
actually_eval <- FALSE
if (actually_eval & focustools::is_monday()) {
  knitr::opts_chunk$set(eval=TRUE)
}
```


# focustools

<!-- badges: start -->
[![R CMD Check (stable)](https://github.com/signaturescience/focustools/workflows/R-CMD-check-stable/badge.svg)](https://github.com/signaturescience/focustools/actions)

[![R CMD Check (dev)](https://github.com/signaturescience/focustools/workflows/R-CMD-check-dev/badge.svg)](https://github.com/signaturescience/focustools/actions)

<!-- badges: end -->

## Installation

Install from GitHub as follows:

``` r
# install.packages("devtools")
devtools::install_github("signaturescience/focustools", build_vignettes = FALSE)
```

## Usage

To get started, see the package vignette:

```{r, eval=FALSE}
vignette("focustools")
```

## COVID-19 Forecast Hub submissions

`focustools` was originally developed to generate forecasts of COVID-19 targets in the United States. The package includes functionality to prepare forecast output for submission to the [COVID-19 Forecast Hub](https://covid19forecasthub.org/). The `focustools` GitHub repository includes a `submission/` directory with select forecast submission files. That same directory tracks a pipeline R script that does the following:

1. Retrieves historical observed count data (national level from JHU by default)
1. Fits incident case and incident death models
1. Forecasts future case data to facilitate the incident death forecast
1. Creates the incident death forecast based on this new data
1. Generates cumulative death forecast by summarizing incident death forecast
1. Prepares submission format
1. Savse the submission to `submission/TEAM-MODEL/YYYY-MM-DD-TEAM-MODEL.csv`
1. Validates the submission file


